---
title: "Wenatchee animal model"
embed-resources: true
format: html
editor: visual
---

## Load and format pedigree and data

```{r}
library(MCMCglmm)
library(ggplot2)
library(sequoia)
library(pedtricks)
library(cowplot)
setwd("/Users/mike.ford/Documents/MikeFord/WenSpChkProp/FRANZWensp_analysis/Git-Wenatchee-RRS")

d = read.csv("pedigree_fitness_Tumwater_Grun_4.29.25.csv",stringsAsFactors = F,na.strings = "NA")
d$id = as.factor(d$id)
d$dam = as.factor(d$dam)
d$sire = as.factor(d$sire)
d = d[,c(3,1,2,4:ncol(d))]
names(d)[1] = "animal"

#### format run timing
d$Date.off = as.Date(d$Date.off,format = '%m/%d/%y')
temp = paste("1/1/",d$Year.off,sep="")
temp = as.Date(temp,format="%m/%d/%Y")
dayofyear = d$Date.off - temp
d$Dayofyear.off = dayofyear
d$Dayofyear.off = as.integer(d$Dayofyear.off)
mean(d$Dayofyear.off,na.rm=T)
sd(d$Dayofyear.off,na.rm=T)

d$Date.sire = as.Date(d$Date.sire,format = '%m/%d/%y')
temp = paste("1/1/",d$Year.sire,sep="")
temp = as.Date(temp,format="%m/%d/%Y")
dayofyear = d$Date.sire - temp
d$Dayofyear.sire = dayofyear
d$Dayofyear.sire = as.integer(d$Dayofyear.sire)
mean(d$Dayofyear.sire,na.rm=T)
sd(d$Dayofyear.sire,na.rm=T)

d$Date.dam = as.Date(d$Date.dam,format = '%m/%d/%y')
temp = paste("1/1/",d$Year.dam,sep="")
temp = as.Date(temp,format="%m/%d/%Y")
dayofyear = d$Date.dam - temp
d$Dayofyear.dam = dayofyear
d$Dayofyear.dam = as.integer(d$Dayofyear.dam)
mean(d$Dayofyear.dam,na.rm=T)
sd(d$Dayofyear.dam,na.rm=T)
```

## Filter pedigree down to the data of interest

```{r}
ped = d
ped = subset(ped,Origin_ped.off == "Hatchery" | Origin_ped.off == "Wild")
ped = subset(ped,!is.na(Origin_ped.off))
ped = subset(ped,!is.na(Origin_ped.dam))
ped = subset(ped,!is.na(Origin_ped.sire))
ped = subset(ped,Sex_Final.off == "Male" | Sex_Final.off == "Female")
ped = subset(ped,GRun.off == "Spring" )
ped = subset(ped,Final_Status_ped.off == "Broodstock" | Final_Status_ped.off == "Spawning Grounds")
ped = subset(ped,Year.off <=2018 & Year.off >= 2008) # after this hatchery fish were not sampled

#ped = subset(ped,Year.off <=2012 & Year.off >= 2008) # or make a smaller pedigree for testing purposes

# limit to neither parent NA
ped = subset(ped, !is.na(sire) & !is.na(dam)) ### try not doing this

```

Sample size per year

```{r}
table(ped$Sex_Field.off,ped$Year.off)
```

Total pedigree size

```{r}
nrow(ped)
```

## Format and complete pedigree

```{r}
pedf = fix_ped(ped)
#dfil = subset(d,animal %in% pedf$id)
dfil = ped

table(dfil$Origin_ped.off,dfil$Year.off)
```

## Fit animal models

## Males only - year as random effect, origin as random effect - fork length

```{r eval = FALSE}
##### set up and run fkl model - 
##### model2 - male only

dmal = subset(dfil,Sex_Final.off=="Male")
pvar = var(dmal$Fork.off,na.rm=T)

priorR2 <- list(
  G = list(G1 = list(V = diag(2)*pvar*.25, nu = .2), G2 = list(V = pvar*.25, nu = 0.2), G3 = list(V=pvar*.25,nu=.2)),
  R = list(R2 = list(V = diag(2)*pvar*.5, nu = .2))
)

# fork length model, age fixed effects and origin and year as random effects
model.Adyo_test_mal <- MCMCglmm(Fork.off ~ as.factor(Age.off) ,
                        random = ~ idh(Origin_ped.off):animal + dam + Year.off ,rcov = ~idh(Origin_ped.off):units,
                        family = "gaussian",
                        pedigree = pedf, data = dmal,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.Adyo_test_mal,file="model1_Adyo_test_mal.Rdata")
```

```{r}
#### load long run from sedna
load("model.Adyo_mal.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_mal$Sol)
```

Plots of random effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_mal$VCV)
summary(model.Adyo_mal)
```

Plot of h2-wild

```{r}
posterior.heritability.fkl_mal_wild <- model.Adyo_mal$VCV[, "Origin_ped.offWild.animal"] /
  (model.Adyo_mal$VCV[, "Origin_ped.offWild.animal"] + model.Adyo_mal$VCV[, "Origin_ped.offWild.units"] + model.Adyo_mal$VCV[, "dam"] + model.Adyo_mal$VCV[, "Year.off"])
plot(posterior.heritability.fkl_mal_wild)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.fkl_mal_wild)
HPDinterval(posterior.heritability.fkl_mal_wild, 0.95)
```

Plot of h2-hatchery

```{r}
posterior.heritability.fkl_mal_hat <- model.Adyo_mal$VCV[, "Origin_ped.offHatchery.animal"] /
  (model.Adyo_mal$VCV[, "Origin_ped.offHatchery.animal"] + model.Adyo_mal$VCV[, "Origin_ped.offHatchery.units"] + model.Adyo_mal$VCV[, "dam"] + model.Adyo_mal$VCV[, "Year.off"])
plot(posterior.heritability.fkl_mal_hat)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.fkl_mal_hat)
HPDinterval(posterior.heritability.fkl_mal_hat, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Adyo_mal$VCV)
```

Parent offspring regression

```{r}
dr = subset(dfil,Sex_Final.off=="Male")
por_males_flk = lm(Fork.off~Fork.sire*Origin_ped.off +as.factor(Age.off),data=dr)
summary(por_males_flk)

testplot_byage = ggplot(data=dr,aes(x=Fork.sire,y=Fork.off,color=as.factor(Age.off),group=Age.off)) + geom_point() + geom_smooth(method=lm) 
testplot_byage

testplot_byage_origin = ggplot(data=dr,aes(x=Fork.sire,y=Fork.off,color=as.factor(Age.off),pch=as.factor(Origin_ped.off))) + geom_point(alpha=.2) + geom_smooth(method=lm) 
testplot_byage_origin

testplot_noage = ggplot(data=dr,aes(x=Fork.sire,y=Fork.off)) + geom_point() + geom_smooth(method=lm) 
testplot_noage

testm = lm(Fork.off~Fork.sire ,data = subset(dr))
summary(testm)
```

## Females only - year as random effect, origin as random effect - fork length

```{r eval = FALSE}
##### set up and run fkl model - 
##### model2 - female only

dmal = subset(dfil,Sex_Final.off=="Female")
pvar = var(dmal$Fork.off,na.rm=T)

priorR2 <- list(
  G = list(G1 = list(V = diag(2)*pvar*.25, nu = .2), G2 = list(V = pvar*.25, nu = 0.2), G3 = list(V=pvar*.25,nu=.2)),
  R = list(R2 = list(V = diag(2)*pvar*.5, nu = .2))
)

# fork length model, age fixed effects and origin and year as random effects
model.Adyo_test_fem <- MCMCglmm(Fork.off ~ as.factor(Age.off) ,
                        random = ~ idh(Origin_ped.off):animal + dam + Year.off ,rcov = ~idh(Origin_ped.off):units,
                        family = "gaussian",
                        pedigree = pedf, data = dmal,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.Adyo_test_fem,file="model1_Adyo_test_fem.Rdata")
```

```{r}
load("model.Adyo_fem.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_fem$Sol)
```

Plots of random effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_fem$VCV)
summary(model.Adyo_fem)
```

Plot of h2-wild

```{r}
posterior.heritability.fkl_fem_wild <- model.Adyo_fem$VCV[, "Origin_ped.offWild.animal"] /
  (model.Adyo_fem$VCV[, "Origin_ped.offWild.animal"] + model.Adyo_fem$VCV[, "Origin_ped.offWild.units"] + model.Adyo_fem$VCV[, "dam"] + model.Adyo_fem$VCV[, "Year.off"])
plot(posterior.heritability.fkl_fem_wild)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.fkl_fem_wild)
HPDinterval(posterior.heritability.fkl_fem_wild, 0.95)
```

Plot of h2-hatchery

```{r}
posterior.heritability.fkl_fem_hat <- model.Adyo_fem$VCV[, "Origin_ped.offHatchery.animal"] /
  (model.Adyo_fem$VCV[, "Origin_ped.offHatchery.animal"] + model.Adyo_fem$VCV[, "Origin_ped.offHatchery.units"] + model.Adyo_fem$VCV[, "dam"] + model.Adyo_fem$VCV[, "Year.off"])
plot(posterior.heritability.fkl_fem_hat)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.fkl_fem_hat)
HPDinterval(posterior.heritability.fkl_fem_hat, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Adyo_fem$VCV)
```

Parent offspring regression

```{r}
dr = subset(dfil,Sex_Final.off=="Female")
por_females_flk = lm(Fork.off~Fork.dam*Origin_ped.off + as.factor(Age.off) ,data=dr)
summary(por_females_flk)
```

## Males only - year as random effect, origin as random effect - run timing

```{r eval = FALSE}
##### set up and run run timing model - 
##### model2 - male only

dmal = subset(dfil,Sex_Final.off=="Male")
pvar = var(dmal$Dayofyear.off,na.rm=T)

priorR2 <- list(
  G = list(G1 = list(V = diag(2)*pvar*.25, nu = .2), G2 = list(V = pvar*.25, nu = 0.2), G3 = list(V=pvar*.25,nu=.2)),
  R = list(R2 = list(V = diag(2)*pvar*.5, nu = .2))
)

#  age fixed effects and origin and year as random effects
model.Adyo_RT_mal <- MCMCglmm(Dayofyear.off ~ as.factor(Age.off) ,
                           random = ~ idh(Origin_ped.off):animal + dam + Year.off ,rcov = ~idh(Origin_ped.off):units,
                           family = "gaussian",
                           pedigree = pedf, data = dmal,
                           nitt = 500000, burnin = 10000, thin = 500,
                           prior = priorR2, verbose = TRUE)
save(model.Adyo_RT_mal,file="model.Adyo_RT_mal.Rdata")
```

```{r}
#### load long run from sedna
load("model.Adyo_RT_mal.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_RT_mal$Sol)
```

Plots of random effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_RT_mal$VCV)
summary(model.Adyo_RT_mal)
```

Plot of h2-wild

```{r}
posterior.heritability.RT_mal_wild <- model.Adyo_RT_mal$VCV[, "Origin_ped.offWild.animal"] /
  (model.Adyo_RT_mal$VCV[, "Origin_ped.offWild.animal"] + model.Adyo_RT_mal$VCV[, "Origin_ped.offWild.units"] + model.Adyo_RT_mal$VCV[, "dam"] + model.Adyo_RT_mal$VCV[, "Year.off"])
plot(posterior.heritability.RT_mal_wild)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.RT_mal_wild)
HPDinterval(posterior.heritability.RT_mal_wild, 0.95)
```

Plot of h2-hatchery

```{r}
posterior.heritability.RT_mal_hat <- model.Adyo_RT_mal$VCV[, "Origin_ped.offHatchery.animal"] /
  (model.Adyo_RT_mal$VCV[, "Origin_ped.offHatchery.animal"] + model.Adyo_RT_mal$VCV[, "Origin_ped.offHatchery.units"] + model.Adyo_RT_mal$VCV[, "dam"] + model.Adyo_RT_mal$VCV[, "Year.off"])
plot(posterior.heritability.RT_mal_hat)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.RT_mal_hat)
HPDinterval(posterior.heritability.RT_mal_hat, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Adyo_RT_mal$VCV)
```

```{r}
dr = subset(dfil,Sex_Final.off=="Male")
por_males_rt = lm(Dayofyear.off~Dayofyear.sire*Origin_ped.off + as.factor(Age.off) ,data=dr)
summary(por_males_rt)
```

## Females only - year as random effect, origin as random effect - run timing

```{r eval = FALSE}
dmal = subset(dfil,Sex_Final.off=="Female")
pvar = var(dmal$Dayofyear.off,na.rm=T)

priorR2 <- list(
  G = list(G1 = list(V = diag(2)*pvar*.25, nu = .2), G2 = list(V = pvar*.25, nu = 0.2), G3 = list(V=pvar*.25,nu=.2)),
  R = list(R2 = list(V = diag(2)*pvar*.5, nu = .2))
)

# fork length model, age fixed effects and origin and year as random effects
model.Adyo_RT_fem <- MCMCglmm(Dayofyear.off ~ as.factor(Age.off) ,
                           random = ~ idh(Origin_ped.off):animal + dam + Year.off ,rcov = ~idh(Origin_ped.off):units,
                           family = "gaussian",
                           pedigree = pedf, data = dmal,
                           nitt = 500000, burnin = 10000, thin = 500,
                           prior = priorR2, verbose = TRUE)
save(model.Adyo_RT_fem,file="model.Adyo_RT_fem.Rdata")
```

```{r}
load("model.Adyo_RT_fem.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_RT_fem$Sol)
```

Plots of random effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_RT_fem$VCV)
summary(model.Adyo_RT_fem)
```

Plot of h2-wild

```{r}
posterior.heritability.RT_fem_wild <- model.Adyo_RT_fem$VCV[, "Origin_ped.offWild.animal"] /
  (model.Adyo_RT_fem$VCV[, "Origin_ped.offWild.animal"] + model.Adyo_RT_fem$VCV[, "Origin_ped.offWild.units"] + model.Adyo_RT_fem$VCV[, "dam"] + model.Adyo_RT_fem$VCV[, "Year.off"])
plot(posterior.heritability.RT_fem_wild)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.RT_fem_wild)
HPDinterval(posterior.heritability.RT_fem_wild, 0.95)
```

Plot of h2-hatchery

```{r}
posterior.heritability.RT_fem_hat <- model.Adyo_RT_fem$VCV[, "Origin_ped.offHatchery.animal"] /
  (model.Adyo_RT_fem$VCV[, "Origin_ped.offHatchery.animal"] + model.Adyo_RT_fem$VCV[, "Origin_ped.offHatchery.units"] + model.Adyo_RT_fem$VCV[, "dam"] + model.Adyo_RT_fem$VCV[, "Year.off"])
plot(posterior.heritability.RT_fem_hat)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.RT_fem_hat)
HPDinterval(posterior.heritability.RT_fem_hat, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Adyo_RT_fem$VCV)
```

Parent offspring regression

```{r}
dr = subset(dfil,Sex_Final.off=="Female")
por_females_rt = lm(Dayofyear.off~Dayofyear.dam*Origin_ped.off + as.factor(Age.off) ,data=dr)
summary(por_females_rt)
```

## Males only - year as random effect, origin as random effect - weight

```{r eval = FALSE}
##### set up and run weight model - 
##### male only

dmal = subset(dfil,Sex_Final.off=="Male")
pvar = var(dmal$Weight.off,na.rm=T)
pvar
priorR2 <- list(
  G = list(G1 = list(V = diag(2)*pvar*.25, nu = .2), G2 = list(V = pvar*.25, nu = 0.2), G3 = list(V=pvar*.25,nu=.2)),
  R = list(R2 = list(V = diag(2)*pvar*.5, nu = .2))
)

# Weight model, age fixed effects and origin and year as random effects
model.Adyo_WT_mal <- MCMCglmm(Weight.off ~ as.factor(Age.off) ,
                           random = ~ idh(Origin_ped.off):animal + dam + Year.off ,rcov = ~idh(Origin_ped.off):units,
                           family = "gaussian",
                           pedigree = pedf, data = dmal,
                           nitt = 500000, burnin = 10000, thin = 500,
                           prior = priorR2, verbose = TRUE)
save(model.Adyo_WT_mal,file="model.Adyo_WT_mal.Rdata")
```

```{r}
#### load long run from sedna
load("model.Adyo_WT_mal.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_WT_mal$Sol)
```

Plots of random effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_WT_mal$VCV)
summary(model.Adyo_WT_mal)
```

Plot of h2-wild

```{r}
posterior.heritability.WT_mal_wild <- model.Adyo_WT_mal$VCV[, "Origin_ped.offWild.animal"] /
  (model.Adyo_WT_mal$VCV[, "Origin_ped.offWild.animal"] + model.Adyo_WT_mal$VCV[, "Origin_ped.offWild.units"] + model.Adyo_WT_mal$VCV[, "dam"] + model.Adyo_WT_mal$VCV[, "Year.off"])
plot(posterior.heritability.WT_mal_wild)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.WT_mal_wild)
HPDinterval(posterior.heritability.WT_mal_wild, 0.95)
```

Plot of h2-hatchery

```{r}
posterior.heritability.WT_mal_hat <- model.Adyo_WT_mal$VCV[, "Origin_ped.offHatchery.animal"] /
  (model.Adyo_WT_mal$VCV[, "Origin_ped.offHatchery.animal"] + model.Adyo_WT_mal$VCV[, "Origin_ped.offHatchery.units"] + model.Adyo_WT_mal$VCV[, "dam"] + model.Adyo_WT_mal$VCV[, "Year.off"])
plot(posterior.heritability.WT_mal_hat)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.WT_mal_hat)
HPDinterval(posterior.heritability.WT_mal_hat, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Adyo_WT_mal$VCV)
```

Parent offspring regression

```{r}
dr = subset(dfil,Sex_Final.off=="Male")
por_males_wt = lm(Weight.off~Weight.sire*Origin_ped.off + as.factor(Age.off) ,data=dr)
summary(por_males_wt)
```

## Females only - year as random effect, origin as random effect - weight

```{r eval = FALSE}
##### set up and run weight model - 
##### female only

dmal = subset(dfil,Sex_Final.off=="Female")
pvar = var(dmal$Weight.off,na.rm=T)
pvar
summary(dmal$Weight.off)
priorR2 <- list(
  G = list(G1 = list(V = diag(2)*pvar*.25, nu = .2), G2 = list(V = pvar*.25, nu = 0.2), G3 = list(V=pvar*.25,nu=.2)),
  R = list(R2 = list(V = diag(2)*pvar*.5, nu = .2))
)

# Weight model, age fixed effects and origin and year as random effects
model.Adyo_WT_fem <- MCMCglmm(Weight.off ~ as.factor(Age.off) ,
                              random = ~ idh(Origin_ped.off):animal + dam + Year.off ,rcov = ~idh(Origin_ped.off):units,
                              family = "gaussian",
                              pedigree = pedf, data = dmal,
                              nitt = 500000, burnin = 10000, thin = 500,
                              prior = priorR2, verbose = TRUE)
save(model.Adyo_WT_fem,file="model.Adyo_WT_fem.Rdata")
```

```{r}
load("model.Adyo_WT_fem.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_WT_fem$Sol)
```

Plots of random effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_WT_fem$VCV)
summary(model.Adyo_WT_fem)
```

Plot of h2-wild

```{r}
posterior.heritability.WT_fem_wild <- model.Adyo_WT_fem$VCV[, "Origin_ped.offWild.animal"] /
  (model.Adyo_WT_fem$VCV[, "Origin_ped.offWild.animal"] + model.Adyo_WT_fem$VCV[, "Origin_ped.offWild.units"] + model.Adyo_WT_fem$VCV[, "dam"] + model.Adyo_WT_fem$VCV[, "Year.off"])
plot(posterior.heritability.WT_fem_wild)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.WT_fem_wild)
HPDinterval(posterior.heritability.WT_fem_wild, 0.95)
```

Plot of h2-hatchery

```{r}
posterior.heritability.WT_fem_hat <- model.Adyo_WT_fem$VCV[, "Origin_ped.offHatchery.animal"] /
  (model.Adyo_WT_fem$VCV[, "Origin_ped.offHatchery.animal"] + model.Adyo_WT_fem$VCV[, "Origin_ped.offHatchery.units"] + model.Adyo_WT_fem$VCV[, "dam"] + model.Adyo_WT_fem$VCV[, "Year.off"])
plot(posterior.heritability.WT_fem_hat)
```

H2 estimate and 95% CI

```{r}
median(posterior.heritability.WT_fem_hat)
HPDinterval(posterior.heritability.WT_fem_hat, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Adyo_WT_fem$VCV)
```

Parent offspring regression

```{r}
dr = subset(dfil,Sex_Final.off=="Female")
por_females_wt = lm(Weight.off~Weight.dam*Origin_ped.off + as.factor(Age.off) ,data=dr)
summary(por_females_wt)
```

## Male age - parent offspring regression

```{r}
dr = subset(dfil,Sex_Final.off=="Male")
dr$Age.off2 = NA
dr$Age.off2 = ifelse(dr$Age.off>=4,0,dr$Age.off2)
dr$Age.off2 = ifelse(dr$Age.off==3,1,dr$Age.off2)

dr$Age.sire2 = NA
dr$Age.sire2 = ifelse(dr$Age.sire>=4,0,dr$Age.sire2)
dr$Age.sire2 = ifelse(dr$Age.sire==3,1,dr$Age.sire2)

table(dr$Age.off2,dr$Age.sire2)

por_males_age = glm(Age.off2~Age.sire2*Origin_ped.off ,data=dr,family = binomial )
summary(por_males_age)
normal_space_coefs = exp(coefficients(por_males_age))/(1 + exp(coefficients(por_males_age)))
normal_space_coefs

### wild only

dr = subset(dfil,Sex_Final.off=="Male" & Origin_ped.off == "Wild")
dr$Age.off2 = NA
dr$Age.off2 = ifelse(dr$Age.off>=4,0,dr$Age.off2)
dr$Age.off2 = ifelse(dr$Age.off==3,1,dr$Age.off2)

dr$Age.sire2 = NA
dr$Age.sire2 = ifelse(dr$Age.sire>=4,0,dr$Age.sire2)
dr$Age.sire2 = ifelse(dr$Age.sire==3,1,dr$Age.sire2)

table(dr$Age.off2,dr$Age.sire2)

por_males_age_wild = glm(Age.off2~Age.sire2 ,data=dr,family = binomial )
summary(por_males_age_wild)
normal_space_coefs_wild = exp(coefficients(por_males_age_wild))/(1 + exp(coefficients(por_males_age_wild)))
normal_space_coefs_wild

### hatchery only

dr = subset(dfil,Sex_Final.off=="Male" & Origin_ped.off == "Hatchery")
dr$Age.off2 = NA
dr$Age.off2 = ifelse(dr$Age.off>=4,0,dr$Age.off2)
dr$Age.off2 = ifelse(dr$Age.off==3,1,dr$Age.off2)

dr$Age.sire2 = NA
dr$Age.sire2 = ifelse(dr$Age.sire>=4,0,dr$Age.sire2)
dr$Age.sire2 = ifelse(dr$Age.sire==3,1,dr$Age.sire2)

table(dr$Age.off2,dr$Age.sire2)

por_males_age_hatchery = glm(Age.off2~Age.sire2 ,data=dr,family = binomial )
summary(por_males_age_hatchery)
normal_space_coefs_hatchery = exp(coefficients(por_males_age_hatchery))/(1 + exp(coefficients(por_males_age_hatchery)))
normal_space_coefs_hatchery


```

## Summary table

```{r echo = F}
st = data.frame(trait = character(12),sex = character(12),origin=character(12),h2=numeric(12),lowerci = numeric(12),upperci = numeric(12))
st$trait = c("FKL","FKL","FKL","FKL","WT","WT","WT","WT","RT","RT","RT","RT")
st$sex = c("male","male","female","female","male","male","female","female","male","male","female","female")
st$origin = c("wild","hatchery","wild","hatchery","wild","hatchery","wild","hatchery","wild","hatchery","wild","hatchery")
st$h2[1] = mean(posterior.heritability.fkl_mal_wild)
st$lowerci[1] = HPDinterval(posterior.heritability.fkl_mal_wild,.95)[1]
st$upperci[1] = HPDinterval(posterior.heritability.fkl_mal_wild)[2]

st$h2[2] = mean(posterior.heritability.fkl_mal_hat)
st$lowerci[2] = HPDinterval(posterior.heritability.fkl_mal_hat,.95)[1]
st$upperci[2] = HPDinterval(posterior.heritability.fkl_mal_hat)[2]

st$h2[3] = mean(posterior.heritability.fkl_fem_wild)
st$lowerci[3] = HPDinterval(posterior.heritability.fkl_fem_wild,.95)[1]
st$upperci[3] = HPDinterval(posterior.heritability.fkl_fem_wild)[2]

st$h2[4] = mean(posterior.heritability.fkl_fem_hat)
st$lowerci[4] = HPDinterval(posterior.heritability.fkl_fem_hat,.95)[1]
st$upperci[4] = HPDinterval(posterior.heritability.fkl_fem_hat)[2]

st$h2[5] = mean(posterior.heritability.WT_mal_wild)
st$lowerci[5] = HPDinterval(posterior.heritability.WT_mal_wild,.95)[1]
st$upperci[5] = HPDinterval(posterior.heritability.WT_mal_wild)[2]

st$h2[6] = mean(posterior.heritability.WT_mal_hat)
st$lowerci[6] = HPDinterval(posterior.heritability.WT_mal_hat,.95)[1]
st$upperci[6] = HPDinterval(posterior.heritability.WT_mal_hat)[2]

st$h2[7] = mean(posterior.heritability.WT_fem_wild)
st$lowerci[7] = HPDinterval(posterior.heritability.WT_fem_wild,.95)[1]
st$upperci[7] = HPDinterval(posterior.heritability.WT_fem_wild)[2]

st$h2[8] = mean(posterior.heritability.WT_fem_hat)
st$lowerci[8] = HPDinterval(posterior.heritability.WT_fem_hat,.95)[1]
st$upperci[8] = HPDinterval(posterior.heritability.WT_fem_hat)[2]

st$h2[9] = mean(posterior.heritability.RT_mal_wild)
st$lowerci[9] = HPDinterval(posterior.heritability.RT_mal_wild,.95)[1]
st$upperci[9] = HPDinterval(posterior.heritability.RT_mal_wild)[2]

st$h2[10] = mean(posterior.heritability.RT_mal_hat)
st$lowerci[10] = HPDinterval(posterior.heritability.RT_mal_hat,.95)[1]
st$upperci[10] = HPDinterval(posterior.heritability.RT_mal_hat)[2]

st$h2[11] = mean(posterior.heritability.RT_fem_wild)
st$lowerci[11] = HPDinterval(posterior.heritability.RT_fem_wild,.95)[1]
st$upperci[11] = HPDinterval(posterior.heritability.RT_fem_wild)[2]

st$h2[12] = mean(posterior.heritability.RT_fem_hat)
st$lowerci[12] = HPDinterval(posterior.heritability.RT_fem_hat,.95)[1]
st$upperci[12] = HPDinterval(posterior.heritability.RT_fem_hat)[2]
st
```

plot h2 results

```{r}
plot = ggplot(data=st,aes(x=origin,y=h2,color=sex,pch=sex,ymin=lowerci,ymax=upperci)) + geom_point() + geom_errorbar(width=.2) + scale_color_manual(values=c("darkseagreen","steelblue"))+ facet_wrap(~trait) + theme_cowplot()
plot
```
