---
title: "Wenatchee animal model"
embed-resources: true
format: html
editor: visual
---

## Load and format pedigree and data

```{r}
library(MCMCglmm)
library(ggplot2)
library(sequoia)
library(pedtricks)
setwd("/Users/mike.ford/Documents/MikeFord/WenSpChkProp/FRANZWensp_analysis/Git-Wenatchee-RRS")

d = read.csv("pedigree_fitness_Tumwater_Grun_4.29.25.csv",stringsAsFactors = F,na.strings = "NA")
d$id = as.factor(d$id)
d$dam = as.factor(d$dam)
d$sire = as.factor(d$sire)
d = d[,c(3,1,2,4:ncol(d))]
names(d)[1] = "animal"
```

## Filter pedigree down to the data of interest

```{r}
ped = d
ped = subset(ped,Origin_ped.off == "Hatchery" | Origin_ped.off == "Wild")
ped = subset(ped,!is.na(Origin_ped.off))
ped = subset(ped,!is.na(Origin_ped.dam))
ped = subset(ped,!is.na(Origin_ped.sire))
ped = subset(ped,Sex_Final.off == "Male" | Sex_Final.off == "Female")
ped = subset(ped,GRun.off == "Spring" )
ped = subset(ped,Final_Status_ped.off == "Broodstock" | Final_Status_ped.off == "Spawning Grounds")
ped = subset(ped,Year.off <=2018 & Year.off >= 2008) # after this hatchery fish were not sampled

#ped = subset(ped,Year.off <=2012 & Year.off >= 2008) # or make a smaller pedigree for testing purposes

# limit to neither parent NA
ped = subset(ped, !is.na(sire) & !is.na(dam)) ### try not doing this

```

Sample size per year

```{r}
table(ped$Sex_Field.off,ped$Year.off)
```

Total pedigree size

```{r}
nrow(ped)
```

## Format and complete pedigree

```{r}
pedf = fix_ped(ped)
#dfil = subset(d,animal %in% pedf$id)
dfil = ped

table(dfil$Origin_ped.off,dfil$Year.off)
```

## Fit animal models

First, try fitting a very simple model on the small data set to see if it works. Will try to estimate the h2 of length. If model already exits, just load.

```{r}

##### set up and run fkl model - 
##### model1 - simple model ignoring sex and origin

priorR2 <- list(
  G = list(G1 = list(V = 1, nu = 0.002), list(V=1,nu=.002)),
  R = list(R2 = list(V = 1, nu = 0.002))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.AdY_test <- MCMCglmm(Fork.off ~ as.factor(Age.off) + as.factor(Year.off) ,
                        random = ~ animal + dam,
                        family = "gaussian",
                        pedigree = pedf, data = dfil,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.AdY_test,file="model1_AdY_test.Rdata")

```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test$VCV)
```

Plot of h2

```{r}
posterior.heritability.fkl <- model.AdY_test$VCV[, "animal"] /
  (model.AdY_test$VCV[, "animal"] + model.AdY_test$VCV[, "units"]+model.AdY_test$VCV[, "dam"] )
plot(posterior.heritability.fkl)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl)
HPDinterval(posterior.heritability.fkl, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.AdY_test$VCV)
```

## See how it works on filtered data - females only

```{r}
##### set up and run fkl model - 
##### model2 - female only

dfem = subset(dfil,Sex_Final.off=="Female")

priorR2 <- list(
  G = list(G1 = list(V = 1, nu = 0.002), list(V=1,nu=.002)),
  R = list(R2 = list(V = 1, nu = 0.002))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.AdY_test_fem <- MCMCglmm(Fork.off ~ as.factor(Age.off) + as.factor(Year.off) ,
                        random = ~ animal + dam,
                        family = "gaussian",
                        pedigree = pedf, data = dfem,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.AdY_test_fem,file="model1_AdY_test_fem.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test_fem$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test_fem$VCV)
```

Plot of h2

```{r}
posterior.heritability.fkl_fem <- model.AdY_test_fem$VCV[, "animal"] /
  (model.AdY_test_fem$VCV[, "animal"] + model.AdY_test_fem$VCV[, "units"] + model.AdY_test_fem$VCV[, "dam"])
plot(posterior.heritability.fkl_fem)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl_fem)
HPDinterval(posterior.heritability.fkl_fem, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.AdY_test_fem$VCV)
```

## Males only

```{r}
##### set up and run fkl model - 
##### model2 - male only

dmal = subset(dfil,Sex_Final.off=="Male")

priorR2 <- list(
  G = list(G1 = list(V = 1, nu = 0.002), list(V=1,nu=.002)),
  R = list(R2 = list(V = 1, nu = 0.002))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.AdY_test_mal <- MCMCglmm(Fork.off ~ as.factor(Age.off) + as.factor(Year.off) ,
                        random = ~ animal + dam,
                        family = "gaussian",
                        pedigree = pedf, data = dmal,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.AdY_test_mal,file="model1_AdY_test_mal.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test_mal$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test_mal$VCV)
```

Plot of h2

```{r}
posterior.heritability.fkl_mal <- model.AdY_test_mal$VCV[, "animal"] /
  (model.AdY_test_mal$VCV[, "animal"] + model.AdY_test_mal$VCV[, "units"] + model.AdY_test_mal$VCV[, "dam"])
plot(posterior.heritability.fkl_mal)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl_mal)
HPDinterval(posterior.heritability.fkl_mal, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.AdY_test_mal$VCV)
```

## Wild Males only

```{r}
##### set up and run fkl model - 
##### model2 - female only

dmalw = subset(dfil,Sex_Final.off=="Male" & Origin_ped.off=="Wild")

priorR2 <- list(
  G = list(G1 = list(V = 1, nu = 0.002), list(V=1,nu=.002)),
  R = list(R2 = list(V = 1, nu = 0.002))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.AdY_test_malw <- MCMCglmm(Fork.off ~ as.factor(Age.off) + as.factor(Year.off) ,
                        random = ~ animal + dam,
                        family = "gaussian",
                        pedigree = pedf, data = dmalw,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.AdY_test_malw,file="model1_AdY_test_malw.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test_malw$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test_malw$VCV)
```

Plot of h2

```{r}
posterior.heritability.fkl_malw <- model.AdY_test_malw$VCV[, "animal"] /
  (model.AdY_test_malw$VCV[, "animal"] + model.AdY_test_malw$VCV[, "units"] + model.AdY_test_malw$VCV[, "dam"])
plot(posterior.heritability.fkl_malw)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl_malw)
HPDinterval(posterior.heritability.fkl_malw, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.AdY_test_malw$VCV)
```

## Hatchery Males only

```{r}
##### set up and run fkl model - 
##### model2 - female only

dmalh = subset(dfil,Sex_Final.off=="Male" & Origin_ped.off=="Hatchery")

priorR2 <- list(
  G = list(G1 = list(V = 1, nu = 0.002), list(V=1,nu=.002)),
  R = list(R2 = list(V = 1, nu = 0.002))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.AdY_test_malh <- MCMCglmm(Fork.off ~ as.factor(Age.off) + as.factor(Year.off) ,
                        random = ~ animal + dam,
                        family = "gaussian",
                        pedigree = pedf, data = dmalh,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.AdY_test_malh,file="model1_AdY_test_malh.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test_malh$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test_malh$VCV)
```

Plot of h2

```{r}
posterior.heritability.fkl_malh <- model.AdY_test_malh$VCV[, "animal"] /
  (model.AdY_test_malh$VCV[, "animal"] + model.AdY_test_malh$VCV[, "units"] + model.AdY_test_malh$VCV[, "dam"] )
plot(posterior.heritability.fkl_malh)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl_malh)
HPDinterval(posterior.heritability.fkl_malh, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.AdY_test_malh$VCV)
```

## Try one without a dam effect

```{r}
dmalw = subset(dfil,Sex_Final.off=="Male" & Origin_ped.off=="Wild")

priorR1 <- list(
  G = list(G1 = list(V = 1, nu = 0.002)),
  R = list(R2 = list(V = 1, nu = 0.002))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.AdY_test_malwnd <- MCMCglmm(Fork.off ~ as.factor(Age.off) + as.factor(Year.off) ,
                        random = ~ animal,
                        family = "gaussian",
                        pedigree = pedf, data = dmalw,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR1, verbose = TRUE)
save(model.AdY_test_malwnd,file="model1_AdY_test_malwnd.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test_malwnd$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AdY_test_malwnd$VCV)
```

Plot of h2

```{r}
posterior.heritability.fkl_malwnd <- model.AdY_test_malwnd$VCV[, "animal"] /
  (model.AdY_test_malwnd$VCV[, "animal"] + model.AdY_test_malwnd$VCV[, "units"]  )
plot(posterior.heritability.fkl_malwnd)
```

Autocorrelation stats

```{r}
autocorr.diag(model.AdY_test_malwnd$VCV)
```

## Sex as a random effect

```{r}
priorR2 <- list(
  G = list(G1 = list(V = diag(2), nu = 1.002), G3 = list(V = 1, nu = 0.002)),
  R = list(R1 = list(V = diag(2), nu = 1.002))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.AsdY_test <- MCMCglmm(Fork.off ~ as.factor(Age.off) + as.factor(Year.off) ,
                        random = ~ idh(Sex_Final.off):animal + dam,
                        rcov = ~ idh(Sex_Final.off):units,
                        family = "gaussian",
                        pedigree = pedf, data = dfil,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.AsdY_test,file="model.AsdY_test.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AsdY_test$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AsdY_test$VCV)
```

Plot of h2 - males

```{r}
summary(model.AsdY_test$VCV)
posterior.heritability.fkl_mal <- model.AsdY_test$VCV[, "Sex_Final.offMale.animal"] /
  (model.AsdY_test$VCV[, "Sex_Final.offMale.animal"] + model.AsdY_test$VCV[, "Sex_Final.offMale.units"] + model.AsdY_test$VCV[, "dam"] )
plot(posterior.heritability.fkl_mal)
```

Plot of h2 - females

```{r}

posterior.heritability.fkl_femal <- model.AsdY_test$VCV[, "Sex_Final.offFemale.animal"] /
  (model.AsdY_test$VCV[, "Sex_Final.offFemale.animal"] + model.AsdY_test$VCV[, "Sex_Final.offFemale.units"] + model.AsdY_test$VCV[, "dam"] )
plot(posterior.heritability.fkl_femal)
```

Autocorrelation stats

```{r}
autocorr.diag(model.AdY_test_malwnd$VCV)
```

## Try year as a random effect

no sex, no origin model

```{r}

##### set up and run fkl model - 
##### model1 - simple model ignoring sex and origin

priorR2 <- list(
  G = list(G1 = list(V = 1, nu = 0.002), G2 = list(V=1,nu=.002), G3 = list(V=1,nu=.002)),
  R = list(R2 = list(V = 1, nu = 0.002))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.Ady_test <- MCMCglmm(Fork.off ~ as.factor(Age.off) ,
                        random = ~ animal + dam + Year.off,
                        family = "gaussian",
                        pedigree = pedf, data = dfil,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.Ady_test,file="model1_Ady_test.Rdata")

```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.Ady_test$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.Ady_test$VCV)
```

Plot of h2

```{r}
posterior.heritability.fkl <- model.Ady_test$VCV[, "animal"] /
  (model.Ady_test$VCV[, "animal"] + model.Ady_test$VCV[, "units"]+model.Ady_test$VCV[, "dam"] + +model.Ady_test$VCV[, "Year.off"] )
plot(posterior.heritability.fkl)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl)
HPDinterval(posterior.heritability.fkl, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Ady_test$VCV)
```

## Females only, year as random effect

```{r}
##### set up and run fkl model - 
##### model2 - female only

dfem = subset(dfil,Sex_Final.off=="Female")

priorR2 <- list(
  G = list(G1 = list(V = 1, nu = 0.002), G2 = list(V=1,nu=.002),G3 = list(V=1,nu=.002)),
  R = list(R2 = list(V = 1, nu = 0.002))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.Ady_test_fem <- MCMCglmm(Fork.off ~ as.factor(Age.off) ,
                        random = ~ animal + dam + Year.off,
                        family = "gaussian",
                        pedigree = pedf, data = dfem,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.Ady_test_fem,file="model1_Ady_test_fem.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.Ady_test_fem$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.Ady_test_fem$VCV)
```

Plot of h2

```{r}
posterior.heritability.fkl_fem <- model.Ady_test_fem$VCV[, "animal"] /
  (model.Ady_test_fem$VCV[, "animal"] + model.Ady_test_fem$VCV[, "units"] + model.Ady_test_fem$VCV[, "dam"] + model.Ady_test_fem$VCV[, "Year.off"])
plot(posterior.heritability.fkl_fem)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl_fem)
HPDinterval(posterior.heritability.fkl_fem, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Ady_test_fem$VCV)
```

## Males only - year as random effect

```{r}
##### set up and run fkl model - 
##### model2 - male only

dmal = subset(dfil,Sex_Final.off=="Male")
pvar = var(dmal$Fork.off,na.rm=T)

priorR2 <- list(
  G = list(G1 = list(V = pvar*.25, nu = 0.2), G2 = list(V=pvar*.25,nu=.2), G3 = list(V=1,nu=.2)),
  R = list(R2 = list(V = pvar*.5, nu = 0.2))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.Ady_test_mal <- MCMCglmm(Fork.off ~ as.factor(Age.off) ,
                        random = ~ animal + dam + Year.off,
                        family = "gaussian",
                        pedigree = pedf, data = dmal,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.Ady_test_mal,file="model1_Ady_test_mal.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.Ady_test_mal$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.Ady_test_mal$VCV)
summary(model.Ady_test_mal$VCV)
```

Plot of h2

```{r}
posterior.heritability.fkl_mal <- model.Ady_test_mal$VCV[, "animal"] /
  (model.Ady_test_mal$VCV[, "animal"] + model.Ady_test_mal$VCV[, "units"] + model.Ady_test_mal$VCV[, "dam"] + model.Ady_test_mal$VCV[, "Year.off"])
plot(posterior.heritability.fkl_mal)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl_mal)
HPDinterval(posterior.heritability.fkl_mal, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Ady_test_mal$VCV)
```

## Males only - year as random effect, origin as random effect - fork length

```{r}
##### set up and run fkl model - 
##### model2 - male only

dmal = subset(dfil,Sex_Final.off=="Male")
pvar = var(dmal$Fork.off,na.rm=T)

priorR2 <- list(
  G = list(G1 = list(V = diag(2)*pvar*.25, nu = .2), G2 = list(V = pvar*.25, nu = 0.2), G3 = list(V=pvar*.25,nu=.2)),
  R = list(R2 = list(V = diag(2)*pvar*.5, nu = .2))
)

# fork length model, age fixed effects and origin and year as random effects
model.Adyo_test_mal <- MCMCglmm(Fork.off ~ as.factor(Age.off) ,
                        random = ~ idh(Origin_ped.off):animal + dam + Year.off ,rcov = ~idh(Origin_ped.off):units,
                        family = "gaussian",
                        pedigree = pedf, data = dmal,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.Adyo_test_mal,file="model1_Adyo_test_mal.Rdata")
```

```{r}
#### load long run from sedna
load("model.Adyo_mal.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_mal$Sol)
```

Plots of random effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_mal$VCV)
summary(model.Adyo_mal$VCV)
```

Plot of h2-wild

```{r}
posterior.heritability.fkl_mal_wild <- model.Adyo_mal$VCV[, "Origin_ped.offWild.animal"] /
  (model.Adyo_mal$VCV[, "Origin_ped.offWild.animal"] + model.Adyo_mal$VCV[, "Origin_ped.offWild.units"] + model.Adyo_mal$VCV[, "dam"] + model.Adyo_mal$VCV[, "Year.off"])
plot(posterior.heritability.fkl_mal_wild)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl_mal_wild)
HPDinterval(posterior.heritability.fkl_mal_wild, 0.95)
```

Plot of h2-hatchery

```{r}
posterior.heritability.fkl_mal_hat <- model.Adyo_mal$VCV[, "Origin_ped.offHatchery.animal"] /
  (model.Adyo_mal$VCV[, "Origin_ped.offHatchery.animal"] + model.Adyo_mal$VCV[, "Origin_ped.offHatchery.units"] + model.Adyo_mal$VCV[, "dam"] + model.Adyo_mal$VCV[, "Year.off"])
plot(posterior.heritability.fkl_mal_hat)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl_mal_hat)
HPDinterval(posterior.heritability.fkl_mal_hat, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Adyo_mal$VCV)
```

## Females only - year as random effect, origin as random effect - fork length

```{r}
##### set up and run fkl model - 
##### model2 - female only

dmal = subset(dfil,Sex_Final.off=="Female")
pvar = var(dmal$Fork.off,na.rm=T)

priorR2 <- list(
  G = list(G1 = list(V = diag(2)*pvar*.25, nu = .2), G2 = list(V = pvar*.25, nu = 0.2), G3 = list(V=pvar*.25,nu=.2)),
  R = list(R2 = list(V = diag(2)*pvar*.5, nu = .2))
)

# fork length model, age fixed effects and origin and year as random effects
model.Adyo_test_fem <- MCMCglmm(Fork.off ~ as.factor(Age.off) ,
                        random = ~ idh(Origin_ped.off):animal + dam + Year.off ,rcov = ~idh(Origin_ped.off):units,
                        family = "gaussian",
                        pedigree = pedf, data = dmal,
                        nitt = 50000, burnin = 1000, thin = 500,
                        prior = priorR2, verbose = TRUE)
save(model.Adyo_test_fem,file="model1_Adyo_test_fem.Rdata")
```

```{r}
load("model.Adyo_fem.Rdata")
```

Plots of fixed effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_fem$Sol)
```

Plots of random effects

```{r}
par(mar=c(3,3,3,3))
plot(model.Adyo_fem$VCV)
summary(model.Adyo_fem$VCV)
```

Plot of h2-wild

```{r}
posterior.heritability.fkl_fem_wild <- model.Adyo_fem$VCV[, "Origin_ped.offWild.animal"] /
  (model.Adyo_fem$VCV[, "Origin_ped.offWild.animal"] + model.Adyo_fem$VCV[, "Origin_ped.offWild.units"] + model.Adyo_fem$VCV[, "dam"] + model.Adyo_fem$VCV[, "Year.off"])
plot(posterior.heritability.fkl_fem_wild)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl_fem_wild)
HPDinterval(posterior.heritability.fkl_fem_wild, 0.95)
```

Plot of h2-hatchery

```{r}
posterior.heritability.fkl_fem_hat <- model.Adyo_fem$VCV[, "Origin_ped.offHatchery.animal"] /
  (model.Adyo_fem$VCV[, "Origin_ped.offHatchery.animal"] + model.Adyo_fem$VCV[, "Origin_ped.offHatchery.units"] + model.Adyo_fem$VCV[, "dam"] + model.Adyo_fem$VCV[, "Year.off"])
plot(posterior.heritability.fkl_fem_hat)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.fkl_fem_hat)
HPDinterval(posterior.heritability.fkl_fem_hat, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Adyo_fem$VCV)
```

## Males only - year as random effect, origin as random effect - age

```{r}
##### set up and run fkl model - 
##### model2 - male only

dmal = subset(dfil,Sex_Final.off=="Male")
dmal$Age.off2 = NA
dmal$Age.off2 = ifelse(dmal$Age.off>=4,0,dmal$Age.off2)
dmal$Age.off2 = ifelse(dmal$Age.off==3,1,dmal$Age.off2)
table(dmal$Age.off2,useNA = 'ifany')


priorB3 <- list(R = list(V = 1, fix = 1),
               G = list(G1 = list(V = diag(2), nu = 1000, alpha.mu = c(0,0), alpha.V = diag(2)), G2 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1),G3 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))

# fork length model, age fixed effects and origin and year as random effects
model.Adyo_age_mal_th <- MCMCglmm(Age.off2 ~ 1,
                        random = ~ idh(Origin_ped.off):animal + dam + Year.off ,
                        family = "threshold",
                        pedigree = pedf, data = dmal,
                        nitt = 500000, burnin = 15000, thin = 500,
                        prior = priorB3, verbose = TRUE)
save(model.Adyo_age_mal_th,file="model.Adyo_age_mal_th.Rdata")
```

```{r}
#### load long run from sedna
load("model.Adyo_age_mal_th.Rdata")
summary(model.Adyo_age_mal_th)
```

Plots of fixed effects

```{r}
#par(mar=c(1,1,1,1))
plot(model.Adyo_age_mal_th$Sol)
summary(model.Adyo_age_mal_th$Sol)
```

Plots of random effects

```{r}
par(mar=c(3,3,3,2))
plot(model.Adyo_age_mal_th$VCV)
summary(model.Adyo_age_mal_th$VCV)
summary(model.Adyo_age_mal_th)
```

```{r}
posterior.heritability.age_mal_wild <- model.Adyo_age_mal_th$VCV[, "Origin_ped.offWild.animal"] /
  (model.Adyo_age_mal_th$VCV[, "Origin_ped.offWild.animal"] + model.Adyo_age_mal_th$VCV[, "Origin_ped.offWild.units"] + model.Adyo_age_mal_th$VCV[, "dam"] + model.Adyo_age_mal_th$VCV[, "Year.off"])
plot(posterior.heritability.age_mal_wild)
  
```

Plot of h2-wild

```{r}
posterior.heritability.age_mal_wild <- model.Adyo_age_mal_th$VCV[, "Origin_ped.offWild.animal"] /
  (model.Adyo_age_mal_th$VCV[, "Origin_ped.offWild.animal"] + model.Adyo_age_mal_th$VCV[, "units"] + model.Adyo_age_mal_th$VCV[, "dam"] + model.Adyo_age_mal_th$VCV[, "Year.off"])
plot(posterior.heritability.age_mal_wild)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.age_mal_wild)
HPDinterval(posterior.heritability.age_mal_wild, 0.95)
```

Plot of h2-hatchery

```{r}
posterior.heritability.age_mal_hat <- model.Adyo_age_mal_th$VCV[, "Origin_ped.offHatchery.animal"] /
  (model.Adyo_age_mal_th$VCV[, "Origin_ped.offHatchery.animal"] + model.Adyo_age_mal_th$VCV[, "Origin_ped.offHatchery.units"] + model.Adyo_age_mal_th$VCV[, "dam"] + model.Adyo_age_mal_th$VCV[, "Year.off"])
plot(posterior.heritability.age_mal_hat)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.age_mal_hat)
HPDinterval(posterior.heritability.age_mal_hat, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Adyo_age_mal_cat$VCV)
```

## Male with origin as random effect but no dam or year

```{r}
##### set up and run fkl model - 
##### model2 - male only

dmal = subset(dfil,Sex_Final.off=="Male")
dmal$Age.off2 = NA
dmal$Age.off2 = ifelse(dmal$Age.off>=4,0,dmal$Age.off2)
dmal$Age.off2 = ifelse(dmal$Age.off==3,1,dmal$Age.off2)



priorB3 <- list(R = list(V = diag(2), fix = 1),
               G = list(G1 = list(V = diag(2), nu = 1000, alpha.mu = c(0,0), alpha.V = diag(2))))


# fork length model, age fixed effects and origin and year as random effects
model.Ao_age_mal <- MCMCglmm(Age.off2 ~ 1,
                        random = ~ idh(Origin_ped.off):animal ,rcov = ~idh(Origin_ped.off):units,
                        family = "threshold",
                        pedigree = pedf, data = dmal,
                        nitt = 500000, burnin = 10000, thin = 500,
                        prior = priorB3, verbose = TRUE)
save(model.Ao_age_mal,file="model.Ao_age_mal.Rdata")
```

```{r}
par(mar=c(3,3,3,3))
plot(model.Ao_age_mal$Sol)
plot(model.Ao_age_mal$VCV)
summary(model.Ao_age_mal)
```

Plot of h2-wild

```{r}
posterior.heritability.age_mal_wild <- model.Ao_age_mal$VCV[, "Origin_ped.offWild.animal"] /
  (model.Ao_age_mal$VCV[, "Origin_ped.offWild.animal"] + model.Ao_age_mal$VCV[, "Origin_ped.offWild.units"] )
plot(posterior.heritability.age_mal_wild)
```

H2 estimate and 95% CI

```{r}
mean(posterior.heritability.age_mal_wild)
HPDinterval(posterior.heritability.age_mal_wild, 0.95)
```

Plot of h2-hatchery

```{r}
posterior.heritability.age_mal_hat <- model.Ao_age_mal$VCV[, "Origin_ped.offHatchery.animal"] /
  (model.Ao_age_mal$VCV[, "Origin_ped.offHatchery.animal"] + model.Ao_age_mal$VCV[, "Origin_ped.offHatchery.units"] )
plot(posterior.heritability.age_mal_hat)
```

H2 estimate and 95% CI

```{r}
mean(posterior.heritability.age_mal_hat)
HPDinterval(posterior.heritability.age_mal_hat, 0.95)
```

Wild only model

```{r}
##### set up and run fkl model - 
##### model2 - wild male only

dmal = subset(dfil,Sex_Final.off=="Male" & Origin_ped.off == "Wild")
dmal$Age.off2 = NA
dmal$Age.off2 = ifelse(dmal$Age.off>=4,0,dmal$Age.off2)
dmal$Age.off2 = ifelse(dmal$Age.off==3,1,dmal$Age.off2)
table(dmal$Age.off2,useNA = 'ifany')
table(dmal$Age.off,dmal$Age.sire)
table(dmal$Origin_ped.off,dmal$Origin_ped.sire)


priorB3 <- list(R = list(V = 1, fix = 1),
               G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1), G2 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1),G3 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))

# fork length model, age fixed effects and origin and year as random effects
model.Adyo_age_mal_th_wild <- MCMCglmm(Age.off2 ~ 1,
                        random = ~ animal + dam + Year.off ,
                        family = "threshold",
                        pedigree = pedf, data = dmal,
                        nitt = 500000, burnin = 15000, thin = 500,
                        prior = priorB3, verbose = TRUE)
save(model.Adyo_age_mal_th_wild,file="model.Adyo_age_mal_th_wild.Rdata")
```

```{r}
plot(model.Adyo_age_mal_th_wild$Sol)
summary(model.Adyo_age_mal_th_wild)
```

```{r}
par(mar=c(3,3,3,2))
plot(model.Adyo_age_mal_th_wild$VCV)
```

Plot of h2-wild

```{r}
posterior.heritability.age_mal_wild <- model.Adyo_age_mal_th_wild$VCV[, "animal"] /
  (model.Adyo_age_mal_th_wild$VCV[, "animal"] + model.Adyo_age_mal_th_wild$VCV[, "units"] + model.Adyo_age_mal_th_wild$VCV[, "dam"] + model.Adyo_age_mal_th_wild$VCV[, "Year.off"])
plot(posterior.heritability.age_mal_wild)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.age_mal_wild)
HPDinterval(posterior.heritability.age_mal_wild, 0.95)
```

Hatchery only model

```{r}
##### set up and run fkl model - 
##### model2 - wild male only

dmal = subset(dfil,Sex_Final.off=="Male" & Origin_ped.off == "Hatchery")
dmal$Age.off2 = NA
dmal$Age.off2 = ifelse(dmal$Age.off>=4,0,dmal$Age.off2)
dmal$Age.off2 = ifelse(dmal$Age.off==3,1,dmal$Age.off2)
table(dmal$Age.off2,useNA = 'ifany')
table(dmal$Age.off,dmal$Age.sire)
table(dmal$Origin_ped.off,dmal$Origin_ped.sire)


priorB3 <- list(R = list(V = 1, fix = 1),
               G = list(G1 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1), G2 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1),G3 = list(V = 1, nu = 1000, alpha.mu = 0, alpha.V = 1)))

# fork length model, age fixed effects and origin and year as random effects
model.Adyo_age_mal_th_hat <- MCMCglmm(Age.off2 ~ 1,
                        random = ~ animal + dam + Year.off ,
                        family = "threshold",
                        pedigree = pedf, data = dmal,
                        nitt = 500000, burnin = 15000, thin = 500,
                        prior = priorB3, verbose = TRUE)
save(model.Adyo_age_mal_th_hat,file="model.Adyo_age_mal_th_hat.Rdata")
```

```{r}
plot(model.Adyo_age_mal_th_hat$Sol)
summary(model.Adyo_age_mal_th_hat)
```

```{r}
par(mar=c(3,3,3,2))
plot(model.Adyo_age_mal_th_hat$VCV)
```

Plot of h2-wild

```{r}
posterior.heritability.age_mal_hat <- model.Adyo_age_mal_th_hat$VCV[, "animal"] /
  (model.Adyo_age_mal_th_hat$VCV[, "animal"] + model.Adyo_age_mal_th_hat$VCV[, "units"] + model.Adyo_age_mal_th_hat$VCV[, "dam"] + model.Adyo_age_mal_th_hat$VCV[, "Year.off"])
plot(posterior.heritability.age_mal_hat)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.age_mal_hat)
HPDinterval(posterior.heritability.age_mal_hat, 0.95)
```

## 

## 

## Plot offspring length and parent length, males

```{r}

lenplot = ggplot(data=subset(dfil,Sex_Final.off=="Male"),aes(x=Fork.sire,y=Fork.off)) + geom_point() + facet_wrap(~Age.off+Age.sire)
lenplot
```
