---
title: "Wenatchee animal model"
embed-resources: true
format: html
editor: visual
---

## Load and format pedigree and data

```{r}
library(MCMCglmm)
library(ggplot2)
library(sequoia)
library(pedtricks)
setwd("/Users/mike.ford/Documents/MikeFord/WenSpChkProp/FRANZWensp_analysis/Git-Wenatchee-RRS")

d = read.csv("pedigree_fitness_Tumwater_Grun_4.29.25.csv",stringsAsFactors = F,na.strings = "NA")
d$id = as.factor(d$id)
d$dam = as.factor(d$dam)
d$sire = as.factor(d$sire)
d = d[,c(3,1,2,4:ncol(d))]
names(d)[1] = "animal"
```

## Filter pedigree down to the data of interest

```{r}
ped = d
ped = subset(ped,Origin_ped.off == "Hatchery" | Origin_ped.off == "Wild")
ped = subset(ped,!is.na(Origin_ped.off))
ped = subset(ped,!is.na(Origin_ped.dam))
ped = subset(ped,!is.na(Origin_ped.sire))
ped = subset(ped,Sex_Final.off == "Male" | Sex_Final.off == "Female")
ped = subset(ped,GRun.off == "Spring" )
ped = subset(ped,Final_Status_ped.off == "Broodstock" | Final_Status_ped.off == "Spawning Grounds")
# limit to neither parent NA
ped = subset(ped, !is.na(sire) & !is.na(dam))

```

Sample size per year

```{r}
table(ped$Final_Status_ped.off,ped$Year.off)
```

Total pedigree size

```{r}
nrow(ped)
```

## Format and complete pedigree

```{r}
pedf = fix_ped(ped)
dfil = subset(d,animal %in% pedf$id)
```

## Fit animal models

First, try fitting a very simple model on the small data set to see if it works. Will try to estimate the h2 of length. If model already exits, just load.

```{r}
load("model_all_two_parent.Rdata")
```

```{r eval = FALSE}

prior1.1 <- list(
  G = list(G1 = list(V = 1, nu = 0.002)),
  R = list(V = 1, nu = 0.002)
)

sml <- MCMCglmm(Fork.off ~ 1,
                     random = ~animal, pedigree = pedf,
                     data = dfil, nitt = 65000, thin = 50, burnin = 15000,
                     prior = prior1.1, verbose = TRUE
)
save(sml,file="model_all_two_parent.Rdata")
```

Plots of fixed effects

```{r}
plot(sml$Sol)
```

Plots of random effects

```{r}
plot(sml$VCV)
```

Plot of h2

```{r}
posterior.heritability.sml <- sml$VCV[, "animal"] /
  (sml$VCV[, "animal"] + sml$VCV[, "units"])
plot(posterior.heritability.sml)
```

H2 estimate and 95% CI

```{r}
posterior.mode(posterior.heritability.sml)
HPDinterval(posterior.heritability.sml, 0.95)
```

Autocorrelation stats

```{r}
autocorr.diag(sml$VCV)
```

## A more complete model (model 2 from Joanna)

load model if it already is run

```{r}
load("model.Asod.Rdata")
```

```{r eval = FALSE}

priorR3 <- list(
  G = list(G1 = list(V = diag(2), nu = 1.002), G2 = list(V = diag(2), nu = 1.002), G3 = list(V = 1, nu = 0.002)),
  R = list(R1 = list(V = diag(2), nu = 1.002), R2 = list(V = diag(2), nu = 1.002))
)

model.Asod <- MCMCglmm(Fork.off ~ Age.off,
                      random = ~ idh(Sex_Final.off):animal + idh(Origin_ped.off):animal + dam,
                      rcov = ~ idh(Sex_Final.off):units + idh(Origin_ped.off):units,
                      family = "gaussian",
                      pedigree = pedf, data = dfil,
                      nitt = 65000, burnin = 1500, thin = 50,
                      prior = priorR3, verbose = TRUE)
save(model.Asod,file="model.Asod.Rdata")
```

Plot fixed effects

```{r}
plot(model.Asod$Sol)
```

Summary of random effects

```{r}
summary(model.Asod$VCV)
```

Autocorrelation stats

```{r}
autocorr.diag(model.Asod$VCV)
```

H2 for hatchery females

```{r}
posterior.heritability.Asod.fem.hat <- (model.Asod$VCV[, "Sex_Final.offFemale.animal"] + model.Asod$VCV[,"Origin_ped.offHatchery.animal"])  /
  (model.Asod$VCV[, "Sex_Final.offFemale.animal"] + model.Asod$VCV[,"Origin_ped.offHatchery.animal"] + model.Asod$VCV[,"dam"] + model.Asod$VCV[,"Sex_Final.offFemale.units"] + model.Asod$VCV[,"Origin_ped.offHatchery.units"] )
plot(posterior.heritability.Asod.fem.hat)
posterior.mode(posterior.heritability.Asod.fem.hat)
HPDinterval(posterior.heritability.Asod.fem.hat, 0.95)
```

H2 for hatchery males

```{r}
posterior.heritability.Asod.male.hat <- (model.Asod$VCV[, "Sex_Final.offMale.animal"] + model.Asod$VCV[,"Origin_ped.offHatchery.animal"])  /
  (model.Asod$VCV[, "Sex_Final.offMale.animal"] + model.Asod$VCV[,"Origin_ped.offHatchery.animal"] + model.Asod$VCV[,"dam"] + model.Asod$VCV[,"Sex_Final.offMale.units"] + model.Asod$VCV[,"Origin_ped.offHatchery.units"] )

plot(posterior.heritability.Asod.male.hat)
posterior.mode(posterior.heritability.Asod.male.hat)
HPDinterval(posterior.heritability.Asod.male.hat, 0.95)
```

H2 for wild females

```{r}
posterior.heritability.Asod.fem.wild <- (model.Asod$VCV[, "Sex_Final.offFemale.animal"] + model.Asod$VCV[,"Origin_ped.offWild.animal"])  /
  (model.Asod$VCV[, "Sex_Final.offFemale.animal"] + model.Asod$VCV[,"Origin_ped.offWild.animal"] + model.Asod$VCV[,"dam"] + model.Asod$VCV[,"Sex_Final.offFemale.units"] + model.Asod$VCV[,"Origin_ped.offWild.units"] )
plot(posterior.heritability.Asod.fem.wild)
posterior.mode(posterior.heritability.Asod.fem.wild)
HPDinterval(posterior.heritability.Asod.fem.wild, 0.95)
```

H2 for wild males

```{r}
posterior.heritability.Asod.male.wild <- (model.Asod$VCV[, "Sex_Final.offMale.animal"] + model.Asod$VCV[,"Origin_ped.offWild.animal"])  /
  (model.Asod$VCV[, "Sex_Final.offMale.animal"] + model.Asod$VCV[,"Origin_ped.offWild.animal"] + model.Asod$VCV[,"Origin_ped.offWild.animal"] + model.Asod$VCV[,"dam"] + model.Asod$VCV[,"Sex_Final.offMale.units"] + model.Asod$VCV[,"Origin_ped.offWild.units"] )
plot(posterior.heritability.Asod.male.wild)
posterior.mode(posterior.heritability.Asod.male.wild)
HPDinterval(posterior.heritability.Asod.male.wild, 0.95)
```
