---
title: "Wenatchee animal model"
embed-resources: true
format: html
editor: visual
---

## Load and format pedigree and data

```{r}
library(MCMCglmm)
library(ggplot2)
library(sequoia)
library(pedtricks)
setwd("/Users/mike.ford/Documents/MikeFord/WenSpChkProp/FRANZWensp_analysis/Git-Wenatchee-RRS")

d = read.csv("pedigree_fitness_Tumwater_Grun_4.29.25.csv",stringsAsFactors = F,na.strings = "NA")
d$id = as.factor(d$id)
d$dam = as.factor(d$dam)
d$sire = as.factor(d$sire)
d = d[,c(3,1,2,4:ncol(d))]
names(d)[1] = "animal"
```

## Filter pedigree down to the data of interest

```{r}
ped = d
ped = subset(ped,Origin_ped.off == "Hatchery" | Origin_ped.off == "Wild")
ped = subset(ped,!is.na(Origin_ped.off))
ped = subset(ped,!is.na(Origin_ped.dam))
ped = subset(ped,!is.na(Origin_ped.sire))
ped = subset(ped,Sex_Final.off == "Male" | Sex_Final.off == "Female")
ped = subset(ped,GRun.off == "Spring" )
ped = subset(ped,Final_Status_ped.off == "Broodstock" | Final_Status_ped.off == "Spawning Grounds")
ped = subset(ped,Year.off <=2018 & Year.off >= 2008) # after this hatchery fish were not sampled
# limit to neither parent NA
ped = subset(ped, !is.na(sire) & !is.na(dam)) ### try not doing this

```

Sample size per year

```{r}
table(ped$Sex_Field.off,ped$Year.off)
```

Total pedigree size

```{r}
nrow(ped)
```

## Format and complete pedigree

```{r}
pedf = fix_ped(ped)
#dfil = subset(d,animal %in% pedf$id)
dfil = ped

table(dfil$Origin_ped.off,dfil$Year.off)
```

## Fit animal models

Models for fork length and run timing were run on sedna (run time approximately two weeks). The code is copied below but is not run in this document.

Fork length

```{r eval = FALSE}
##### set up and run fkl model

priorR3 <- list(
  G = list(G1 = list(V = diag(2), nu = 1.002), G2 = list(V = diag(2), nu = 1.002), G3 = list(V = 1, nu = 0.002)),
  R = list(R1 = list(V = diag(2), nu = 1.002), R2 = list(V = diag(2), nu = 1.002))
)

# fork length model, age and year fixed effects and sex and origin as random effects
model.AsodY <- MCMCglmm(Fork.off ~ as.factor(Age.off) + as.factor(Year.off) ,
                        random = ~ idh(Sex_Final.off):animal + idh(Origin_ped.off):animal + dam,
                        rcov = ~ idh(Sex_Final.off):units + idh(Origin_ped.off):units,
                        family = "gaussian",
                        pedigree = pedf, data = dfil,
                        nitt = 500000, burnin = 1000, thin = 500,
                        prior = priorR3, verbose = TRUE)
save(model.AsodY,file="model.AsodY.A_as_factor_Y_as_factor.Rdata")

```

Load model output here, copied from sedna

```{r}
load("model.AsodY.A_as_factor_Y_as_factor.Rdata")
```

Autocorrelation stats

```{r}
autocorr.diag(model.AsodY$VCV)
```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AsodY$Sol)
summary(model.AsodY$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.AsodY$VCV)
summary(model.AsodY$VCV)
```

Calculate and plot of h2

Hatchery-origin males

```{r}
h2.fkl.hat.male <- (model.AsodY$VCV[, "Sex_Final.offMale.animal"] + model.AsodY$VCV[,"Origin_ped.offHatchery.animal"]) /
  (model.AsodY$VCV[, "Sex_Final.offMale.animal"] + model.AsodY$VCV[,"Origin_ped.offHatchery.animal"] + model.AsodY$VCV[, "Sex_Final.offMale.units"] + model.AsodY$VCV[,"Origin_ped.offHatchery.units"]  + model.AsodY$VCV[, "dam"])
plot(h2.fkl.hat.male)
summary(h2.fkl.hat.male)
```

H2 estimate and 95% CI

```{r}
posterior.mode(h2.fkl.hat.male)
HPDinterval(h2.fkl.hat.male, 0.95)
```

Natural-origin males

```{r}
h2.fkl.nat.male <- (model.AsodY$VCV[, "Sex_Final.offMale.animal"] + model.AsodY$VCV[,"Origin_ped.offWild.animal"]) /
  (model.AsodY$VCV[, "Sex_Final.offMale.animal"] + model.AsodY$VCV[,"Origin_ped.offWild.animal"] + model.AsodY$VCV[, "Sex_Final.offMale.units"] + model.AsodY$VCV[,"Origin_ped.offWild.units"]  + model.AsodY$VCV[, "dam"])
plot(h2.fkl.nat.male)
summary(h2.fkl.nat.male)
```

H2 estimate and 95% CI

```{r}
posterior.mode(h2.fkl.nat.male)
HPDinterval(h2.fkl.nat.male, 0.95)
```

Hatchery-origin females

```{r}
h2.fkl.hat.female <- (model.AsodY$VCV[, "Sex_Final.offFemale.animal"] + model.AsodY$VCV[,"Origin_ped.offHatchery.animal"]) /
  (model.AsodY$VCV[, "Sex_Final.offFemale.animal"] + model.AsodY$VCV[,"Origin_ped.offHatchery.animal"] + model.AsodY$VCV[, "Sex_Final.offFemale.units"] + model.AsodY$VCV[,"Origin_ped.offHatchery.units"]  + model.AsodY$VCV[, "dam"])
plot(h2.fkl.hat.female)

```

H2 estimate and 95% CI

```{r}
posterior.mode(h2.fkl.hat.female)
HPDinterval(h2.fkl.hat.female, 0.95)
```

Natural-origin females

```{r}
h2.fkl.nat.female <- (model.AsodY$VCV[, "Sex_Final.offFemale.animal"] + model.AsodY$VCV[,"Origin_ped.offWild.animal"]) /
  (model.AsodY$VCV[, "Sex_Final.offFemale.animal"] + model.AsodY$VCV[,"Origin_ped.offWild.animal"] + model.AsodY$VCV[, "Sex_Final.offFemale.units"] + model.AsodY$VCV[,"Origin_ped.offWild.units"]  + model.AsodY$VCV[, "dam"])
plot(h2.fkl.nat.female)
summary(h2.fkl.nat.female)
```

H2 estimate and 95% CI

```{r}
posterior.mode(h2.fkl.nat.female)
HPDinterval(h2.fkl.nat.female, 0.95)
```

Run timing

```{r eval = FALSE}
priorR3 <- list(
  G = list(G1 = list(V = diag(2), nu = 1.002), G2 = list(V = diag(2), nu = 1.002), G3 = list(V = 1, nu = 0.002)),
  R = list(R1 = list(V = diag(2), nu = 1.002), R2 = list(V = diag(2), nu = 1.002))
)

# run timing model
model.rt.AsodY <- MCMCglmm(Dayofyear.off ~ as.factor(Age.off) + as.factor(Year.off) ,
                        random = ~ idh(Sex_Final.off):animal + idh(Origin_ped.off):animal + dam,
                        rcov = ~ idh(Sex_Final.off):units + idh(Origin_ped.off):units,
                        family = "gaussian",
                        pedigree = pedf, data = dfil,
                        nitt = 500000, burnin = 1000, thin = 500,
                        prior = priorR3, verbose = TRUE)
save(model.rt.AsodY,file="model.rt.AsodY.A_as_factor_Y_as_factor.Rdata")

```

Load model output here, copied from sedna

```{r}
load("model.rt.AsodY.A_as_factor_Y_as_factor.Rdata")
```

Autocorrelation stats

```{r}
autocorr.diag(model.rt.AsodY$VCV)
```

Plots of fixed effects

```{r}
par(mar=c(1,1,1,1))
plot(model.rt.AsodY$Sol)
summary(model.rt.AsodY$Sol)
```

Plots of random effects

```{r}
par(mar=c(1,1,1,1))
plot(model.rt.AsodY$VCV)
summary(model.rt.AsodY$VCV)
```

Calculate and plot of h2

Hatchery-origin males

```{r}
h2.rt.hat.male <- (model.rt.AsodY$VCV[, "Sex_Final.offMale.animal"] + model.rt.AsodY$VCV[,"Origin_ped.offHatchery.animal"]) /
  (model.rt.AsodY$VCV[, "Sex_Final.offMale.animal"] + model.rt.AsodY$VCV[,"Origin_ped.offHatchery.animal"] + model.rt.AsodY$VCV[, "Sex_Final.offMale.units"] + model.rt.AsodY$VCV[,"Origin_ped.offHatchery.units"]  + model.rt.AsodY$VCV[, "dam"])
plot(h2.rt.hat.male)
summary(h2.rt.hat.male)
```

H2 estimate and 95% CI

```{r}
posterior.mode(h2.rt.hat.male)
HPDinterval(h2.rt.hat.male, 0.95)
```

Natural-origin males

```{r}
h2.rt.nat.male <- (model.rt.AsodY$VCV[, "Sex_Final.offMale.animal"] + model.rt.AsodY$VCV[,"Origin_ped.offWild.animal"]) /
  (model.rt.AsodY$VCV[, "Sex_Final.offMale.animal"] + model.rt.AsodY$VCV[,"Origin_ped.offWild.animal"] + model.rt.AsodY$VCV[, "Sex_Final.offMale.units"] + model.rt.AsodY$VCV[,"Origin_ped.offWild.units"]  + model.rt.AsodY$VCV[, "dam"])
plot(h2.rt.nat.male)
summary(h2.rt.nat.male)
```

H2 estimate and 95% CI

```{r}
posterior.mode(h2.rt.nat.male)
HPDinterval(h2.rt.nat.male, 0.95)
```

Hatchery-origin females

```{r}
h2.rt.hat.female <- (model.rt.AsodY$VCV[, "Sex_Final.offFemale.animal"] + model.rt.AsodY$VCV[,"Origin_ped.offHatchery.animal"]) /
  (model.rt.AsodY$VCV[, "Sex_Final.offFemale.animal"] + model.rt.AsodY$VCV[,"Origin_ped.offHatchery.animal"] + model.rt.AsodY$VCV[, "Sex_Final.offFemale.units"] + model.rt.AsodY$VCV[,"Origin_ped.offHatchery.units"]  + model.rt.AsodY$VCV[, "dam"])
plot(h2.rt.hat.female)

```

H2 estimate and 95% CI

```{r}
posterior.mode(h2.rt.hat.female)
HPDinterval(h2.rt.hat.female, 0.95)
```

Natural-origin females

```{r}
h2.rt.nat.female <- (model.rt.AsodY$VCV[, "Sex_Final.offFemale.animal"] + model.rt.AsodY$VCV[,"Origin_ped.offWild.animal"]) /
  (model.rt.AsodY$VCV[, "Sex_Final.offFemale.animal"] + model.rt.AsodY$VCV[,"Origin_ped.offWild.animal"] + model.rt.AsodY$VCV[, "Sex_Final.offFemale.units"] + model.rt.AsodY$VCV[,"Origin_ped.offWild.units"]  + model.rt.AsodY$VCV[, "dam"])
plot(h2.rt.nat.female)
summary(h2.rt.nat.female)
```

H2 estimate and 95% CI

```{r}
posterior.mode(h2.rt.nat.female)
HPDinterval(h2.rt.nat.female, 0.95)
```
