---
title: "Wenatchee parent offspring regression"
embed-resources: true
format: html
editor: visual
---

## Load and format pedigree and data

```{r}
library(MCMCglmm)
library(ggplot2)
library(sequoia)
library(pedtricks)
library(cowplot)
setwd("/Users/mike.ford/Documents/MikeFord/WenSpChkProp/FRANZWensp_analysis/Git-Wenatchee-RRS")

d = read.csv("pedigree_fitness_Tumwater_Grun_4.29.25.csv",stringsAsFactors = F,na.strings = "NA")
d$id = as.factor(d$id)
d$dam = as.factor(d$dam)
d$sire = as.factor(d$sire)
d = d[,c(3,1,2,4:ncol(d))]
names(d)[1] = "animal"

#### format run timing
d$Date.off = as.Date(d$Date.off,format = '%m/%d/%y')
temp = paste("1/1/",d$Year.off,sep="")
temp = as.Date(temp,format="%m/%d/%Y")
dayofyear = d$Date.off - temp
d$Dayofyear.off = dayofyear
d$Dayofyear.off = as.integer(d$Dayofyear.off)
mean(d$Dayofyear.off,na.rm=T)
sd(d$Dayofyear.off,na.rm=T)

d$Date.sire = as.Date(d$Date.sire,format = '%m/%d/%y')
temp = paste("1/1/",d$Year.sire,sep="")
temp = as.Date(temp,format="%m/%d/%Y")
dayofyear = d$Date.sire - temp
d$Dayofyear.sire = dayofyear
d$Dayofyear.sire = as.integer(d$Dayofyear.sire)
mean(d$Dayofyear.sire,na.rm=T)
sd(d$Dayofyear.sire,na.rm=T)

d$Date.dam = as.Date(d$Date.dam,format = '%m/%d/%y')
temp = paste("1/1/",d$Year.dam,sep="")
temp = as.Date(temp,format="%m/%d/%Y")
dayofyear = d$Date.dam - temp
d$Dayofyear.dam = dayofyear
d$Dayofyear.dam = as.integer(d$Dayofyear.dam)
mean(d$Dayofyear.dam,na.rm=T)
sd(d$Dayofyear.dam,na.rm=T)
```

## Filter pedigree down to the data of interest

```{r}
ped = d
ped = subset(ped,Origin_ped.off == "Hatchery" | Origin_ped.off == "Wild")
ped = subset(ped,!is.na(Origin_ped.off))
ped = subset(ped,!is.na(Origin_ped.dam))
ped = subset(ped,!is.na(Origin_ped.sire))
ped = subset(ped,Sex_Final.off == "Male" | Sex_Final.off == "Female")
ped = subset(ped,GRun.off == "Spring" )
ped = subset(ped,Final_Status_ped.off == "Broodstock" | Final_Status_ped.off == "Spawning Grounds")
ped = subset(ped,Year.off <=2018 & Year.off >= 2008) # after this hatchery fish were not sampled

#ped = subset(ped,Year.off <=2012 & Year.off >= 2008) # or make a smaller pedigree for testing purposes

# limit to neither parent NA
ped = subset(ped, !is.na(sire) & !is.na(dam)) ### 

```

Sample size per year

```{r}
table(ped$Sex_Field.off,ped$Year.off)
```

Total pedigree size

```{r}
nrow(ped)
```

## Parent offspring regression for various traits

The goal of this exercise is a) to visualize the relationships between parent and offspring traits for various subsets of the data set, and b) provide a comparison to the heritability estimates from the animal models. One specific goal is to try to understand why some of the animal model heritability estimates are so high.

## Fork length

Parent offspring regression -- males, regression to mid-parent, including offspring age as a factor

```{r}
dr = ped
dr$Fork.mid = (dr$Fork.dam+dr$Fork.sire)/2
summary(dr$Fork.mid)
rflk_males_age = lm(Fork.off~Fork.mid*Origin_ped.off +as.factor(Age.off),data=subset(dr,Sex_Final.off == "Male"))
summary(rflk_males_age)

rflk_male_age_plot = ggplot(data=subset(dr,Sex_Final.off == "Male"),aes(x=Fork.mid,y=Fork.off,color=as.factor(Age.off),group=Age.off)) + geom_point(alpha=.2) + geom_smooth(method=lm) 
rflk_male_age_plot


```

Males, regression to sire, offspring age as factor

```{r}
dr = ped
rflk_males_age_sire = lm(Fork.off~Fork.sire*Origin_ped.off +as.factor(Age.off),data=subset(dr,Sex_Final.off == "Male"))
summary(rflk_males_age_sire)

rflk_male_age_sire_plot = ggplot(data=subset(dr,Sex_Final.off == "Male"),aes(x=Fork.sire,y=Fork.off,color=as.factor(Age.off),group=Age.off)) + geom_point(alpha=.2) + geom_smooth(method=lm) 
rflk_male_age_sire_plot
```

Males, regression to sire, offspring age AND sire age as factors (no origin)

```{r}
dr = ped
dr$Agei = paste(dr$Age.sire,dr$Age.off,sep="_")
table(dr$Agei)
rflk_males_age2_sire = lm(Fork.off~Fork.sire+as.factor(Age.off)+as.factor(Age.sire),data=subset(dr,Sex_Final.off == "Male"))
summary(rflk_males_age2_sire)

rflk_male_age2_sire_plot = ggplot(data=subset(dr,Sex_Final.off == "Male"),aes(x=Fork.sire,y=Fork.off,color=as.factor(Agei),group=Agei)) + geom_point(alpha=.2) + geom_smooth(method=lm) 
rflk_male_age2_sire_plot
```

Males, offspring and sire are age 4 (most common age)

```{r}
dr = ped

rflk_males_4 = lm(Fork.off~Fork.sire*Origin_ped.off,data=subset(dr,Sex_Final.off == "Male" & Age.off==4 & Age.sire==4))
summary(rflk_males_4)

rflk_male_4_plot = ggplot(data=subset(dr,Sex_Final.off == "Male" & Age.off==4 & Age.sire==4),aes(x=Fork.sire,y=Fork.off,color=Origin_ped.off)) + geom_point(alpha=.2) + geom_smooth(method=lm) 
rflk_male_4_plot
```

Males, offspring and sire are both age 3

```{r}
dr = ped

rflk_males_3 = lm(Fork.off~Fork.sire*Origin_ped.off,data=subset(dr,Sex_Final.off == "Male" & Age.off==3 & Age.sire==3))
summary(rflk_males_3)

rflk_male_3_plot = ggplot(data=subset(dr,Sex_Final.off == "Male" & Age.off==3 & Age.sire==3),aes(x=Fork.sire,y=Fork.off,color=Origin_ped.off)) + geom_point(alpha=.2) + geom_smooth(method=lm) 
rflk_male_3_plot
```

Males, offspring and sire are both age 5

```{r}
dr = ped

rflk_males_5 = lm(Fork.off~Fork.sire*Origin_ped.off,data=subset(dr,Sex_Final.off == "Male" & Age.off==5 & Age.sire==5))
summary(rflk_males_5)

rflk_male_5_plot = ggplot(data=subset(dr,Sex_Final.off == "Male" & Age.off==5 & Age.sire==5),aes(x=Fork.sire,y=Fork.off,color=Origin_ped.off)) + geom_point(alpha=.2) + geom_smooth(method=lm) 
rflk_male_5_plot
```

Females, regresssion to mid-parent

```{r}
dr = ped
dr$Fork.mid = (dr$Fork.dam+dr$Fork.sire)/2
summary(dr$Fork.mid)
rflk_females_age = lm(Fork.off~Fork.mid*Origin_ped.off +as.factor(Age.off),data=subset(dr,Sex_Final.off == "Female"))
summary(rflk_females_age)

rflk_female_age_plot = ggplot(data=subset(dr,Sex_Final.off == "Female"),aes(x=Fork.mid,y=Fork.off,color=as.factor(Age.off),group=Age.off)) + geom_point(alpha=.2) + geom_smooth(method=lm) 
rflk_female_age_plot
```

Females, regression to dam

```{r}
dr = ped
dr$Fork.mid = (dr$Fork.dam+dr$Fork.sire)/2
summary(dr$Fork.mid)
rflk_females_age_dam = lm(Fork.off~Fork.dam*Origin_ped.off +as.factor(Age.off),data=subset(dr,Sex_Final.off == "Female"))
summary(rflk_females_age_dam)

rflk_female_age_dam_plot = ggplot(data=subset(dr,Sex_Final.off == "Female"),aes(x=Fork.dam,y=Fork.off,color=as.factor(Age.off),group=Age.off)) + geom_point(alpha=.2) + geom_smooth(method=lm) 
rflk_female_age_dam_plot
```

## Run timing

Males, mid-parent, including year as factor since it varies so among years

```{r}
dr = ped
dr$Dayofyear.mid = (dr$Dayofyear.dam+dr$Dayofyear.sire)/2
summary(dr$Dayofyear.mid)
rRT_males = lm(Dayofyear.off~Dayofyear.mid*Origin_ped.off +as.factor(Year.off),data=subset(dr,Sex_Final.off == "Male"))
summary(rRT_males)

rRT_male_plot = ggplot(data=subset(dr,Sex_Final.off == "Male"),aes(x=Dayofyear.mid,y=Dayofyear.off,color=as.factor(Year.off),pch=Origin_ped.off)) + geom_point(alpha=.2) + geom_smooth(method=lm) 
rRT_male_plot
```

Females

```{r}
dr = ped
dr$Dayofyear.mid = (dr$Dayofyear.dam+dr$Dayofyear.sire)/2
summary(dr$Dayofyear.mid)
rRT_females = lm(Dayofyear.off~Dayofyear.mid*Origin_ped.off +as.factor(Year.off),data=subset(dr,Sex_Final.off == "Female"))
summary(rRT_females)

rRT_female_plot = ggplot(data=subset(dr,Sex_Final.off == "Female"),aes(x=Dayofyear.mid,y=Dayofyear.off,color=as.factor(Year.off),pch=Origin_ped.off)) + geom_point(alpha=.2) + geom_smooth(method=lm) 
rRT_female_plot
```

## Age

Males. Will do this several ways. First, see determine if there is a relationship between offspring and sire age via contingency table analyses

```{r}
dr = ped
maleagetable = table(dr$Age.off,dr$Age.sire)
maleagetable
mosaicplot(maleagetable,xlab="Offspring Age",ylab="Sire age")
```

```{r}
maleagechisq = chisq.test(maleagetable,simulate.p.value = T)
round(maleagechisq$observed - maleagechisq$expected,2)

```

Age appears to be heritable - a sire is more likely than expected to produce a male offspring of the same age (excluded age 2). Let's see if this differs between hatchery and wild

```{r}
dr = ped
drwild = subset(dr,Origin_ped.off == "Wild")
drhatchery = subset(dr,Origin_ped.off =="Hatchery")
maleagetablewild = table(drwild$Age.off,drwild$Age.sire)
maleagetablewild
maleagetablehatchery = table(drhatchery$Age.off,drhatchery$Age.sire)
maleagetablehatchery

mosaicplot(maleagetablewild,xlab="Offspring Age",ylab="Sire age",main="Wild")
mosaicplot(maleagetablehatchery,xlab="Offspring Age",ylab="Sire age",main="Hatchery")

```

```{r}
maleagechisqwild = chisq.test(maleagetablewild,simulate.p.value = T)
print("wild")
round((maleagechisqwild$observed - maleagechisqwild$expected),2)

maleagechisqhatchery = chisq.test(maleagetablehatchery,simulate.p.value = T)
print("hatchery")
round((maleagechisqhatchery$observed - maleagechisqhatchery$expected),2)


```

```{r}
ageplotvio = ggplot(data = subset(ped,Sex_Final.off=="Male"),aes(x=Age.sire,y=as.factor(Age.off))) + geom_violin()
ageplotvio
```

```{r}
ageplotreg = ggplot(data = subset(ped,Sex_Final.off=="Male"),aes(x=Age.sire,y=Age.off)) + geom_jitter() + geom_smooth(method=lm)
ageplotreg
```

```{r}

```

Try treating age as just a normal trait

```{r}
dr = ped
rage_males = lm(Age.off~Age.sire*Origin_ped.off ,data=subset(dr,Sex_Final.off == "Male"))
summary(rage_males)

ageplotreg = ggplot(data = subset(ped,Sex_Final.off=="Male"),aes(x=Age.sire,y=Age.off)) + geom_jitter() + geom_smooth(method=lm)
ageplotreg
```

Note, could try binomial or multinomial regression, but I find interpreting the results from those models to be very difficult.

## Fitness (offspring number)

Males

```{r}
dr = ped
dr$noff.mid = (dr$noff.dam+dr$noff.sire)/2
summary(dr$noff.mid)
rfit_males = lm(noff.off~noff.mid*Origin_ped.off +as.factor(Year.off)+Final_Status_ped.off,data=subset(dr,Sex_Final.off == "Male"))
summary(rfit_males)

rfit_male_plot = ggplot(data=subset(dr,Sex_Final.off == "Male"),aes(x=noff.mid,y=noff.off,color=as.factor(Year.off),pch=Origin_ped.off)) + geom_point(alpha=.2) + geom_smooth(method=lm)  + facet_wrap(vars(Final_Status_ped.off,Final_Status_ped.dam),scales = "free")
rfit_male_plot
```
