---
title: "Wenatchee RS analyses"

format: html
embed-resources: true
editor: visual
---

## Preliminary analysis of Wenatchee River Chinook reproductive success data

Michael Ford 2.17.25

```{r}
library(ggplot2)

ped = read.csv("/Users/mike.ford/Documents/MikeFord/WenSpChkProp/FRANZWensp_analysis/pedigree_fitness_Tumwater_Grun_3.12.25.csv",stringsAsFactors = F)

### do some basic summaries
pedf = subset(ped,Origin.off == "Hatchery" | Origin.off == "Wild")
pedf = subset(pedf,!is.na(Origin.off))
pedf = subset(pedf,Sex_Final.off == "Male" | Sex_Final.off == "Female")
table(pedf$GRun.off)
pedf = subset(pedf,GRun.off == "Spring" )
table(pedf$Final_Status.off)

pedf = subset(pedf,Final_Status.off == "Broodstock" | Final_Status.off == "Spawning Grounds")
table(pedf$Final_Status.off)
```

### Summarize the number of location and origin calls that we changed due to location/origin contradictions

Location changes

```{r}
table(pedf$Final_Status.off,pedf$Final_Status_ped.off)
```

Origin Changes

```{r}
table(pedf$Origin.off,pedf$Origin_ped.off)
```

### Annual counts by sex and origin

```{r}
#counts per year by sex and origin
#pdf("year counts by sex and origin.pdf")
ggplot(data=pedf,aes(Year.off,fill=Origin_ped.off)) + geom_bar() + facet_wrap(~Sex_Final.off) + scale_fill_manual(values=c("darkseagreen","steelblue"))
#dev.off()
```

### Annual run timing distributions

```{r}
pedf$Date.off = as.Date(pedf$Date.off,format = '%m/%d/%y')
temp = paste("1/1/",pedf$Year.off,sep="")
temp = as.Date(temp,format="%m/%d/%Y")
dayofyear = pedf$Date.off - temp
pedf$Dayofyear.off = dayofyear

#pdf("run time by year.pdf")
ggplot(data=pedf,aes(y=as.factor(Year.off),x=Dayofyear.off)) + geom_violin(draw_quantiles = .5)
#dev.off()


```

### Annual length (POH) distribution, by sex

```{r}
#pdf("POH by year and sex.pdf")
ggplot(data=pedf,aes(y=as.factor(Year.off),x=POH.off,fill=Sex_Final.off)) + geom_violin(draw_quantiles = .5) + scale_fill_manual(values=c("darkseagreen","steelblue"))
#dev.off()
```

### Annual length (POH) distribution by origin, original

```{r}
#pdf("POH by year and origin.pdf")
ggplot(data=pedf,aes(y=as.factor(Year.off),x=POH.off,fill=Origin.off)) + geom_violin(draw_quantiles = .5) + scale_fill_manual(values=c("darkseagreen","steelblue")) 
#dev.off()
```

### Annual length (POH) distribution by origin, pedigree corrected

```{r}
#pdf("POH by year and origin.pdf")
ggplot(data=pedf,aes(y=as.factor(Year.off),x=POH.off,fill=Origin_ped.off)) + geom_violin(draw_quantiles = .5) + scale_fill_manual(values=c("darkseagreen","steelblue")) 
#dev.off()
```

### 

### Dot plot of reproductive success (RS = adult offspring number) by location

```{r}
# RS by year, origin, sex, and location
#pdf("RS by year and origin and final location.pdf")
ggplot(data=subset(pedf),aes(y=as.factor(Year.off),x=noff.off,color=Origin_ped.off)) + geom_jitter() + scale_color_manual(values=c("darkseagreen","steelblue")) + facet_wrap(~Final_Status_ped.off,scales="free")
#dev.off()
```

```{r}
#### make a basic RS and RRS tables, limited to comparisons of interest
makeRStable = function(d)
{
  rst = by(d,d[,c("Year.off")],FUN=function(x) t.test(noff.off~Origin_ped.off,data=x))
  rst
  rsn = by(d,d[,c("Year.off","Origin_ped.off")],FUN=function(x) nrow(x))
  dimnames(rsn)
  Nhat = rsn[,1]
  Nwild = rsn[,2]
  
  years = unlist(dimnames(rst)[1])
  hatmeans = unlist(lapply(rst,function(x) x$estimate[1]))
  wildmeans = unlist(lapply(rst,function(x) x$estimate[2]))
  pvals = unlist(lapply(rst,function(x) x$p.value))
  rrs = hatmeans/wildmeans
  basicrrs = data.frame(Year=years, Hatchery_N = Nhat, Hatchery_RS = hatmeans, Wild_N = Nwild, Wild_RS=wildmeans, RRS = rrs, P_value=pvals,row.names = NULL)
  return(basicrrs)
}
```

### Table of RS and RRS by year, for males in the stream environment

```{r}
d = subset(pedf,Year.off <= 2018 & Final_Status.off=="Spawning Grounds" & Sex_Final.off=="Male")
RS.male.stream = makeRStable(d)
RS.male.stream
write.csv(RS.male.stream,"RRS table males stream origin_ped.csv")
```

### Table of RS and RRS by year, for females in the stream environment

```{r}
d = subset(pedf,Year.off <= 2018 & Final_Status.off=="Spawning Grounds" & Sex_Final.off=="Female")
RS.female.stream = makeRStable(d)
RS.female.stream
write.csv(RS.female.stream,"RRS table females stream origin_ped.csv")
```

### Table of RS and RRS by year, for males in the broodstock

```{r}
d = subset(pedf,Year.off <= 2013 & Final_Status.off=="Broodstock" & Sex_Final.off=="Male")
RS.male.brood = makeRStable(d)
RS.male.brood
write.csv(RS.male.brood,"RRS males brood origin_ped.csv")
```

### Table of RS and RRS by year, for females in the broodstock

```{r}
d = subset(pedf,Year.off <= 2013 & Final_Status.off=="Broodstock" & Sex_Final.off=="Female")
RS.female.brood = makeRStable(d)
RS.female.brood
write.csv(RS.female.brood,"RRS females brood origin_ped.csv")
```

```{r}
#### look at the effects of parental origin (=grandparent spawning location) on RS
# first create a new column that provides parental origin
pedf$Origin.parents = ifelse(!is.na(pedf$Origin_ped.dam) & !is.na(pedf$Origin_ped.sire),paste(pedf$Origin_ped.dam,pedf$Origin_ped.sire,sep="-"),NA)

makeRSGEN2table = function(d)
{
  rst = by(d,d[,c("Year.off")],FUN=function(x) summary(lm(noff.off~Origin.parents,data=x)))
  rst[[1]]$coefficients
  rsn = by(d,d[,c("Year.off","Origin.parents")],FUN=function(x) nrow(x))
  dimnames(rsn)
  Nhathat = rsn[,1]
  Nhatwild = rsn[,2]
  Nwildhat = rsn[,3]
  Nwildwild = rsn[,4]
  
  years = unlist(dimnames(rst)[1])
  hathatmeans = unlist(lapply(rst,function(x) x$coefficients[1,1]))
  hatwildmeans = unlist(lapply(rst,function(x) x$coefficients[2,1])) + hathatmeans
  wildhatmeans = unlist(lapply(rst,function(x) x$coefficients[3,1])) + hathatmeans
  wildwildmeans = unlist(lapply(rst,function(x) x$coefficients[4,1])) + hathatmeans
  pvalhatwild = unlist(lapply(rst,function(x) x$coefficients[2,4]))
  pvalwildhat = unlist(lapply(rst,function(x) x$coefficients[3,4]))
  pvalwildwild = unlist(lapply(rst,function(x) x$coefficients[4,4]))
  rrsHW = hatwildmeans/hathatmeans
  rrsWH = wildhatmeans/hathatmeans
  rrsWW = wildwildmeans/hathatmeans
  basicrrs = data.frame(Year=years, HH_N = Nhathat,HW_N=Nhatwild,WH_N=Nwildhat,WW_N=Nwildwild,HH_RS=hathatmeans, HW_RS = hatwildmeans,WH_RS=wildhatmeans,WW_RS=wildwildmeans, RRS_HW = rrsHW,RRS_WH=rrsWH,RRS_WW=rrsWW, P_value_HW=pvalhatwild,P_value_WH=pvalwildhat,P_value_WW=pvalwildwild,row.names = NULL)
  return(basicrrs)
}
```

### Annual table for 2 generation RRS in the stream environment

```{r}
d = subset(pedf,Year.off <= 2018 & Final_Status_ped.off=="Spawning Grounds" & Final_Status_ped.dam=="Spawning Grounds" &  Final_Status.sire=="Spawning Grounds" )
RSgen2.both.stream = makeRSGEN2table(d)
RSgen2.both.stream
```

### 2 generation RRS in the stream environment, all years combined

```{r}
d = subset(pedf,Year.off <= 2018 & Final_Status.off=="Spawning Grounds" & Final_Status.dam=="Spawning Grounds" &  Final_Status.sire=="Spawning Grounds")
RSgen2.both.stream.allyears = lm(noff.off~Origin.parents,data=d)
summary(RSgen2.both.stream.allyears)
```

### 2 generation RRS in the broodstock , all years combined

```{r}
d = subset(pedf,Year.off <= 2013 & Final_Status.off=="Broodstock" & Final_Status.dam=="Broodstock" &  Final_Status.sire=="Broodstock" )
RSgen2.both.brood.allyears = lm(noff.off~Origin.parents,data=d)
summary(RSgen2.both.brood.allyears)
```

### Effects of number of wild grandparents on RRS

```{r}
### note - this file was created using the field-only origin - not ped-corrected
anc = read.csv("Ancestor_origins_1.22.25.csv",stringsAsFactors = F)
pedf = merge(pedf,anc,by="id",all.x=T)
```

### Both sexes, stream environment

```{r}
# both sexes, spawning grounds
d = subset(pedf,Year.off <= 2018 & Final_Status.off=="Spawning Grounds" & N.wild.gp>0)
GPlm = lm(noff.off~as.factor(N.wild.gp),data=d)
summary(GPlm)
```

```{r}
# males, spawning grounds
d = subset(pedf,Year.off <= 2018 & Final_Status.off=="Spawning Grounds" & N.wild.gp>0 & Sex_Final.off=="Male")
GPlm.male = lm(noff.off~as.factor(N.wild.gp),data=d)
summary(GPlm.male)
```

```{r}
# females, spawning grounds
d = subset(pedf,Year.off <= 2018 & Final_Status.off=="Spawning Grounds" & N.wild.gp>0 & Sex_Final.off=="Female")
GPlm.female = lm(noff.off~as.factor(N.wild.gp),data=d)
summary(GPlm.female)
```

### In the broodstock...

```{r}
# both, brood
d = subset(pedf,Year.off <= 2013 & Final_Status.off=="Broodstock" & N.wild.gp>0)
GPlmbrood = lm(noff.off~as.factor(N.wild.gp),data=d)
summary(GPlmbrood)
```

```{r}
# males, brood
d = subset(pedf,Year.off <= 2013 & Final_Status.off=="Broodstock" & N.wild.gp>0 & Sex_Final.off=="Male")
GPlmbrood.male = lm(noff.off~as.factor(N.wild.gp),data=d)
summary(GPlmbrood.male)
```

```{r}
# females, brood
d = subset(pedf,Year.off <= 2013 & Final_Status.off=="Broodstock" & N.wild.gp>0 & Sex_Final.off=="Female")
GPlmbrood.female = lm(noff.off~as.factor(N.wild.gp),data=d)
summary(GPlmbrood.female)
```

### Graphs of relationships between RS and traits

Male run timing, stream environment

```{r}
d = subset(pedf,Year.off <= 2018 & Final_Status_ped.off=="Spawning Grounds" & Sex_Final.off=="Male")
rt.plot.males.stream = ggplot(data=d,aes(x=Dayofyear.off,y=noff.off))  + geom_smooth(method="loess") + facet_wrap(~Year.off) 
rt.plot.males.stream
```

Male POH, stream environment

```{r}
POH.plot.males.stream = ggplot(data=d,aes(x=POH.off,y=noff.off))  + geom_smooth(method="loess") + facet_wrap(~Year.off)
POH.plot.males.stream
```

```{r}
d = subset(pedf,Year.off <= 2013 & Final_Status_ped.off=="Broodstock" & Sex_Final.off=="Male")
rt.plot.males.brood = ggplot(data=d,aes(x=Dayofyear.off,y=noff.off))  + geom_smooth(method="loess") + facet_wrap(~Year.off)
POH.plot.males.brood = ggplot(data=d,aes(x=POH.off,y=noff.off))  + geom_smooth(method="loess") + facet_wrap(~Year.off)
```

Male run timing, in broodstock

```{r}
rt.plot.males.brood
```

Male POH, in broostock

```{r}
POH.plot.males.brood
```

```{r}
d = subset(pedf,Year.off <= 2018 & Final_Status_ped.off=="Spawning Grounds" & Sex_Final.off=="Female")
rt.plot.females.stream = ggplot(data=d,aes(x=Dayofyear.off,y=noff.off))  + geom_smooth(method="loess") + facet_wrap(~Year.off) 
POH.plot.females.stream = ggplot(data=d,aes(x=POH.off,y=noff.off))  + geom_smooth(method="loess") + facet_wrap(~Year.off)
```

Female run timing, stream environment

```{r}
rt.plot.females.stream
```

Female POH, stream environment

```{r}
POH.plot.females.stream
```

```{r}
d = subset(pedf,Year.off <= 2013 & Final_Status_ped.off=="Broodstock" & Sex_Final.off=="Female")
rt.plot.females.brood = ggplot(data=d,aes(x=Dayofyear.off,y=noff.off))  + geom_smooth(method="loess") + facet_wrap(~Year.off)
POH.plot.females.brood = ggplot(data=d,aes(x=POH.off,y=noff.off))  + geom_smooth(method="loess") + facet_wrap(~Year.off)
```

Female run timing, broodstock

```{r}
rt.plot.females.brood
```

Female POH, broostock

```{r}
POH.plot.females.brood
```
